Alternative Perspective of the RSA Cipher

### 2017-05-17 latex/keyexchange/secretkeybasedpublickeyencryption.tex
added new cipher idea

\section{Introduction}
 The ability to exchange cryptographic keys over insecure channels is of vital importance to the function of the internet; Without algorithms for arranging secure communication online, much of every day life in the modern world would cease to be possible.

The introduction of Quantum Computing and post-quantum cryptography has lead to the development of lattice based cryptographic schemes that are believed to be secure even when faced with quantum parallelism. Simply put, RSA and DH will need replacement at some point, and lattices seem like a promising alternative.

@@

Where the $r$ values are of appropriate size relative to $s_1$ and $s_2$ \todo{what sizes are appropriate}
While decryption is computed as:

@@

\subsection{Public Key Randomization}
Because the public key consists of malleable homomorphic ciphertexts, it can be randomized without complicating any of the rest of the procedures. The presented scheme can utilize encrypted public keys, and we give two potential use cases for the idea.

>>>>

\section{Introduction}
 The ability to exchange cryptographic keys over insecure channels is of vital importance for secure communications in general; Without algorithms for arranging secure communication online, much of every day life in the modern world would cease to be possible.

The introduction of Quantum Computing and post-quantum cryptography has lead to the development of lattice based cryptographic schemes that are believed to be secure even when faced with quantum parallelism.

@@

Where the $r$ values are of appropriate size relative to $s_1$ and $s_2$.

Decryption is computed as:

@@

\subsection{Public Key Randomization}
\todo{Needs more development}
Because the public key consists of malleable homomorphic ciphertexts, it can be randomized without complicating any of the rest of the procedures. The presented scheme can utilize encrypted public keys, and we give two potential use cases for the idea.

### 2017-05-17 latex/rsareviewed.tex
added new cipher idea

\title{Brief Article}
\author{The Author}
%\date{} % Activate to display a given date or no date (if empty),
         % otherwise the current date is printed

@@

\section{Cipher Constructions}

The design of a cipher can be split into (at least) two parts: The cipher construction, and the core permutation. The two archetypal examples of cipher construction are the Feistel Network and Substiution-Permutation Network.

In the former, the state is divided into two sections that are typically of equal size; A key is added to some side(s), then a core "randomizing" function is applied to one half to produce a key stream; This key stream is then combined with the other half, in order to encipher it. We can see then, that the core "randomizing function" is interchangeable within the Feistel Network construction.

In a Substitution-Permutation network, the state is typically subdivided into many smaller sections. The algorithm typically consists of addition of a key layer, followed by at least one application of an invertible "randomizing" function applied to the entire state; This is repeated until a sufficient "security margin" is obtained. This construction can be modeled as the composition of two functions: The key addition function, and the application of the core permutation on the state. As in the Feistel Network, the core permutation is interchangeable within the Substitution-Permutation network construction.

We can see then, that there are at least two, effectively separate aspects to cipher design: The cipher construction, and the core permutation. This work focuses on the former. While the core permutation is critical for resistance to generic attacks such as linear and differential cryptanalysis, there also exist generic attacks against the cipher constructions. Additionally, while the design of the permutation can be hugely influential on the performance of the algorithm, so too can the cipher construction.

\section{Construction Design Considerations}

First, the target platform needs to be established. What is ideal for constrained hardware is not necessarily ideal for a modern CPU. In addition to the target platform(s), the performance goals of the algorithm should be outlined: Is latency important, or is throughput for large messages important? Is it important to use as little (or much) RAM?

Less obvious concerns that are still very important include: Is resistance to side channel leakage minimized? Does it facilitate or impede attempts to implement counter-measures?

"The simple" - RC4 like (shuffling huge array of words)

"The small" - 8-bit friendly/constrained device friendly.

"The fast" - salsa like (SIMD with big registers)

\subsection{A subsection}

More text.

>>>>

\title{Alternative Perspective of the RSA Cipher}
\author{Ella Rose}
%\date{} % Activate to display a given date or no date (if empty),
         % otherwise the current date is printed

@@

\begin{abstract}
 We provide an alternative description of the RSA cipher by breaking it up into a few discrete, separate pieces. We hope that by understanding RSA as part of a more generalized cipher framework, we  might come up with alternative instantiations of the cipher that utilize different hardness assumptions. We do not suggest such a design here.

\section{Introduction to RSA}
This section will outline the RSA algorithm - for the history of the design, we encourage the reader to look elsewhere. \todo{RSA history is where?}

The RSA cryptosystem utilizes three core functions:

 The key generation algorithm, which generates a private key and corresponding public key
 The encryption algorithm, which replaces the supplied number with a psuedorandom number
 The decryption algorithm, which recovers the number that was supplied to the encryption algorithm, when given the output of the encryption algorithm and the private key.

The key generation algorithm is typically described by something like the following:

Select two large, random prime numbers of approximately equal size $p, q \gets bigPrime(size) bigPrime(size)$
Calculate the totient via $totient \gets (p -1)(q - 1)$
Set e to the usual/safe value (i.e. 65537); Select the decryption exponent via $modularInverse(e, totient)
Multiply p and q to produce the modulus n $n \gets pq$
Output n, e as the public key; Output d as the public key;

While the encryption/decryption algorithms are simplier and are described by something like:

Produce a ciphertext of message by computing $message ^ e \mod n$
Produce a plaintext from ciphertext by computing $ciphertext ^ d mod n$

While this is enough information to produce a working implementation of the "textbook" RSA algorithm, it is does not really fit well into a generic cipher framework.

\section{An Alternative Introduction to RSA}
First, we shall describe a simple secret key cipher that is homomorphic with respect to multiplication of ciphertexts:

Select a random prime number of sufficient size $p \gets bigPrime(size)$
Select an encryption exponent of appropriate size (>2 and <size of p)
Calculate the decryption exponent by computing $d \gets modularInverse(e, p - 1)$
Encrypt a number by computing $m^e mod n$ and decrypt by computing $c^d mod n$

The only difference so far is that we are using a single prime number as the modulus. Despite the fact that encryption and decryption are asymmetric operations, this cipher is not secure as a public key cipher, as anyone who knows $p$ and $e$ may compute $d$.

RSA typically computes the decryption exponent by computing $modularInverse(e, totient(n))$, but it works equally well to use $modularInverse(e, totient(p))$ or $modularInverse(e, totient(q))$ - as long as the message is smaller then p or q, respectively.

So the modulus $n \gets np$ that is typically released as part of the public key is effectively an encrypted modulus.

As long as we encrypt messages of size smaller then p (the basis), then we do not need to know the totient of n - we only need to know the totient of p (or at least, one factor of n).

# the purpose of q is to encrypt p so that others cannot obtain d
# mk is a secure cipher if m and k are large primes and no "fast" classical algorithm for factoring exists
